---
title: "ES sites in Pakistan - 2021-24"
output:
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: 72
---

# Background

Confidence in the interruption of poliovirus transmission is highly
dependent on assumed sensitivity of environmental surveillance (ES), in
particular in the case of cVDPV2 (low rate of clinical disease).
Previous work has explored how the strength of evidence against ongoing
circulation is influenced by variation in which sites (in which
locations, relative to the distribution of population) are reporting
samples at what time. This work also incorporated detection rate of
NPEVs as another source of variation between sites, hence variation in
the strength of evidence for absence in that location when no poliovirus
has been detected.

The latter indicator was considered only with respect to a fixed target
rate defined by the GPEI (\>50%), which is used to monitor site
performance and inform their opening/closure. This means that sites
detecting NPEVs in an average of 50% of collected samples over the
previous 12 months are considered as performing at their "maximum", with
our measurement of sensitivity scaled down when this value falls below
50%. In fact, the 50% threshold is arbitrarily defined and doesn't
necessarily align with the expected prevalence of NPEVs in the
population, in different locations and at different times of year.

In this setting, the nature of collection sites varies considerably
between natural waterways, informal drainage channels, and more formal
sewage networks. These different types of site will likely have very
different population catchment and virus detection sensitivity. Sites
are, however, only defined by name and GPS location. In order to
distinguish different site types, we can look at the land cover
surrounding the site and the proximity/intersection with natural water
bodies.

Ultimately, we want to infer the relationship between probability of
NPEV detection and: - Population size within Xkm - Dominant land cover
type within Xkm - Proximity to natural waterways/bodies - Elevation
within Xkm compared to elevation at site? [- Proportion of "flooded
vegetation" within Xkm]

For the last point, we are interested in the influence of seasonal
flooding on NPEV detection. This may alter the usual catchment of a
site, as well as diluting/contaminating samples with run-off from other
regions. However, this is not a significant consideration in the context
of Nigeria and should instead be explored in a more flood-prone setting
such as Pakistan.

The WHO establishes ES sites to supposedly capture \~100,000 people,
however in reality this often isn't the case. Previous work has
demonstrated significant inconsistency in the interpretation of site
catchment. Different sites will capture different-sized areas according
to the nature of the site, therefore it is not clear what catchment
radius is appropriate to assumed for analysis. This could be
investigated by

Methodological approach:

-   Fit GAM of P(NPEV) \~ catch_pop + lc_type + proximity_water -
    Investigate anisotropy in the correlation structure via directional
    variograms.

-   Investigate shift in Pr NPEV detection during July/August 2022,
    comparing sites within/without flood-affected area.

## Set up environment and load data

```{r}

pacman::p_load(tidyverse, sf, janitor, terra, exactextractr, mgcv, gstat, broom)
theme_set(theme_minimal())

# Inputs
indir <- "data/Pakistan"
proj_local <- "EPSG:32631"

# ES linelist
es <- readRDS(file.path(indir,"es_analysis.rds")) |> 
  mutate(nonpolio = is.na(final_class)|final_class %in% c("NPEV","VACCINE"))

# AFP linelist
afp <- readRDS(file.path(indir,"afp_analysis.rds"))

# Population raster
poprast  <- rast(file.path(indir,"proj_pop_rast.tif"))

# Administrative boundaries
shape2 <- readRDS(file.path(indir, "shape2.rds")) |> 
  st_transform(proj_local)

shape0 <- st_union(shape2)

```

# Descriptive

## Unique sites and characteristics

```{r}

es |> 
  select(guid, admin_0:admin_2, site_id, site_name, x, y) |> 
  unique() -> es_sites

glimpse(es_sites)

```

In total, `r nrow(es_sites)` unique sites active at some point between
`r min(es$collection_date)` and `r max(es$collection_date)`.

Check uniqueness of coordinates per site code:

```{r}

nrow(es_sites)
n_distinct(es_sites$site_id)

```

```{r}

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_as_sf(es_sites, coords = c("x","y"), crs = 4326), 
          pch = 22, fill = "white") + 
  labs(title = "Unique wastewater sampling sites in Pakistan",
       caption = "At least one sample collected Jan 2021-Dec 2024.")

```

### Site names

Can we glean anything from site names? - Some in particular refer to
bridges, specific streams/rivers, community hubs (e.g.
school/mosque/camp/police station) and road intersections
(junction/roundabout) - There is one formal wastewater treatment plant

What proportion of samples from identifiable site types are
NPEV-positive?

```{r}

es_sites |> 
  mutate(keywords = gsub("[ ]", "", site_name)) |> separate_rows(keywords, sep = " ") |>
  group_by(keywords) |> 
  summarise(count = n()) |> 
  arrange(desc(count))

```

```{r}

es |> 
  mutate(site_name = tolower(site_name),
         drain = grepl("drain",site_name) | grepl("sewage",site_name) | 
           grepl("disposal",site_name) | grepl("sewer",site_name) ,
         water = grepl("stream",site_name) | grepl("river",site_name) | 
           grepl("pond",site_name) | grepl("canal",site_name),
         comm_site = grepl("school",site_name) | grepl("college",site_name) | 
           grepl("market",site_name) | grepl("masjid",site_name),
         hosp = grepl("hospital", site_name),
         junction = grepl("junction",site_name),
         wwtp = (grepl("pumping",site_name)|
                   (grepl("pump", site_name) & grepl("station", site_name))|
                   grepl("plant", site_name)),
         no_type = !(drain|water|comm_site|hosp|junction|wwtp)) -> tmp

tmp |> summarise(across(drain:no_type, sum)) 

tmp |> 
  pivot_longer(drain:no_type, names_to = "type") |> 
  filter(value, type != "no_type") |> 
  ggplot(aes(type, fill = !is.na(npev))) + 
  geom_bar() + 
  labs(fill = "NPEV+", x = NULL, y = "Count", title = "Sample count and positivity by rough site type")

```

Not really helpful...

Just distinguish WWTPs:

```{r}

es_sites |> 
  full_join(select(tmp, site_id, wwtp) |> distinct()) |> 
  mutate(wwtp_f = factor(wwtp, levels = c(FALSE,TRUE), labels = c("Other","PS or TP"))) |> 
  arrange(wwtp_f) -> es_sites

tabyl(es_sites$wwtp)

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data =  st_as_sf(es_sites, coords = c("x","y"), crs = 4326), 
          aes(fill = wwtp_f),
          pch = 22) + 
  scale_fill_manual("Site type",values = c("white","blue1")) +
  labs(title = "Unique wastewater sampling sites in Pakistan",
       subtitle = "Pumping stations and sewage treatment plants",
       caption = "At least one sample collected Jan 2021-Dec 2024.")

```

Estimate crude detection rates by site type:

```{r}

tmp |> 
  group_by(wwtp) |>
  summarise(n = n(),
            npev_pos = sum(!is.na(npev)),
            lo = Hmisc::binconf(npev_pos, n)[2],
            hi = Hmisc::binconf(npev_pos, n)[3]) |>
  ungroup() |> 
  mutate(wwtp_f = factor(wwtp, levels = c(FALSE,TRUE), labels = c("Other","PS or TP"))) |> 
  ggplot(aes(wwtp_f, npev_pos/n, ymin = lo, ymax = hi)) + 
  geom_errorbar(width = 0.2) +
  geom_point() +
  theme(legend.position = c(0.1,0.1)) + 
  labs(fill = "NPEV+", 
       x = NULL, y = "Proportion positive", 
       title = "Sample positivity by site type",
       subtitle = "95% binomial confidence intervals")

tmp |> 
  filter(nonpolio) |> 
  group_by(wwtp) |>
  summarise(n = n(),
            npev_pos = sum(!is.na(npev)),
            lo = Hmisc::binconf(npev_pos, n)[2],
            hi = Hmisc::binconf(npev_pos, n)[3]) |>
  ungroup() |> 
  mutate(grp = "Excluding PV+") -> tmp2

tmp |> 
  group_by(wwtp) |>
  summarise(n = n(),
            npev_pos = sum(!is.na(npev)),
            lo = Hmisc::binconf(npev_pos, n)[2],
            hi = Hmisc::binconf(npev_pos, n)[3]) |>
  ungroup() |> 
  mutate(grp = "Including PV+") |> 
  bind_rows(tmp2) |> 
  mutate(wwtp_f = factor(wwtp, levels = c(FALSE,TRUE), labels = c("Other","PS or TP"))) |> 
  ggplot(aes(wwtp_f, npev_pos/n, ymin = lo, ymax = hi, lty = grp)) + 
  geom_errorbar(width = 0.2) +
  geom_point() +
  scale_linetype_manual(values = c("dashed",'solid')) +
  theme(legend.position = c(0.1,0.1)) + 
  labs(fill = "NPEV+", 
       lty = NULL,
       x = NULL, y = "Proportion positive", 
       title = "Sample positivity by site type",
       subtitle = "95% binomial confidence intervals")

```

### Catchment population

```{r}

es_sites_utm <- st_as_sf(es_sites, coords = c("x","y"), crs = 4326) |> 
  st_transform(proj_local)

es_buff <- st_buffer(es_sites_utm, dist = units::set_units(2, "km"))

# Population within 2km of each site:
es_sites_utm$catchment_2k <- exactextractr::exact_extract(poprast, es_buff, "sum") 

# Total population per admin2 area:
shape2$total_pop <- exactextractr::exact_extract(poprast, shape2, "sum")
    
hist(es_sites_utm$catchment_2k)
summary(es_sites_utm$catchment_2k)

```

On average, `r mean(es_sites_utm$catchment_2k, na.rm = T)` people reside
within 2km of an individual collection site, with a total of
`r sum(es_sites_utm$catchment_2k)` within 2km of *any* site
(\~`r round(sum(es_sites_utm$catchment_2k)/sum(shape2$total_pop)*100)`%
of the national population).

```{r}

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = es_sites_utm |> st_transform(4326), aes(size = catchment_2k, colour = catchment_2k), pch = 15, alpha = 0.7) + 
  scale_colour_viridis_c(trans = "sqrt", labels = scales::label_comma()) +
  theme(legend.position = c(0.9,0.25)) + 
  guides(size = "none") + 
  labs(colour = NULL,
       title = "Catchment population within 2km of site")

```

### Sample frequency

How many samples are collected per site?

```{r}

es |> 
  mutate(collection_month = month,
         month = lubridate::month(month, label = T)) |> 
  group_by(collection_month, month, year, site_id, x, y) |> 
  summarise(
    n_samples = n(),
    n_cvdpv2 = sum(final_class == "cVDPV2", na.rm = T),
    n_npev = sum(!is.na(npev))) |> 
  ungroup() -> es_mth
  
tabyl(es_mth, n_samples)

tabyl(es_mth, month, n_samples)

ggplot(st_drop_geometry(es_mth)) + 
  geom_bar(aes(month, fill = as.factor(n_samples))) + 
  labs(y = "Number of sites", x = "Month", fill = "Sampling\nfrequency")

```

Fewer sites sampling twice-monthly during summer?

The majority of sites are reporting one sample per month, with the rest
reporting two. On very few occasions has a site reported more than two
samples in a given month. Which sites are these? Do some consistently
report more than others?

```{r}

es_mth |> 
  filter(n_samples > 2) |> 
  st_as_sf(coords = c("x","y"), crs = 4326) |> 
  st_jitter() |> 
  ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(pch = "+", cex = 5, alpha = 0.5) + 
  theme_void()

```

### Sample quality

Reported sample quality likely links to the sensitivity for NPEV
detection, however few samples are reported as something other than
"good":

```{r}

tabyl(es$sample_condition)
ggplot(es, aes(year, fill = sample_condition)) + 
  geom_bar()

es |> 
  group_by(site_id, x, y) |> 
  summarise(p_qual = mean(!is.na(sample_condition) & sample_condition == "Good")) |> 
  ungroup() |> 
  st_as_sf(coords = c("x","y"), crs = 4326) -> tmp
ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = tmp, aes(fill = p_qual), pch = 22) + 
  scale_fill_viridis_c(limits = c(0,1)) +
  theme(legend.position = c(0.9,0.25)) + 
  labs(fill = NULL,
       title = "Overall proportion of good-condition samples",
       caption = "At least one sample collected Jan 2021-Dec 2024.")

```

### Land cover within catchment area

If we can't really distinguish site types from their name, perhaps we
can infer something from the type environment surrounding the site
location, e.g. - Urbanicity - Land cover - Proximity to water

Load rasters of land use and distance to waterways: [Note that these
from WorldPop are quite out of date - should be replaced with ESRI land
cover estimates]

```{r}

# water_dist  <- rast("data/nga_osm_dst_waterway_100m_2016.tif") |> 
#   terra::project(proj_local)
# 
# built_extent  <- rast("data/nga_bsgme_v0a_100m_2020.tif") |> 
#   terra::project(proj_local)

```

```{r}
# 
# plot(water_dist)
# plot(built_extent)

```

-   Built extent raster is binary, but seems to be aggregated to a
    different resolution in autoplot hence non-binary values.
-   Water distance raster is in km to the nearest *major* waterway.

First look at the point value at each site (create a small buffer to
still use exact_extract):

```{r}

# es_buff_sm <- st_buffer(es_sites_utm, dist = units::set_units(50, "m"))
# 
# es_sites_utm$built_extent <- exactextractr::exact_extract(built_extent, es_buff_sm, "mean")
# es_sites_utm$water_dist <- exactextractr::exact_extract(water_dist, es_buff_sm, "mean") 
# 
# summary(es_sites_utm$built_extent)
# hist(es_sites_utm$built_extent)
# 
# summary(es_sites_utm$water_dist)
# hist(es_sites_utm$water_dist)

```

Also look at average value within 2km catchment:

```{r}

# es_sites_utm$built_2k <- exactextractr::exact_extract(built_extent, 
#                                                       es_buff,
#                                                       "mean") 
# es_sites_utm$water_2k <- exactextractr::exact_extract(water_dist, 
#                                                       es_buff,
#                                                       "mean") 
# 
# ggplot(es_sites_utm, aes(built_extent, built_2k)) + geom_abline() + geom_point(alpha = 0.5) 
# ggplot(es_sites_utm, aes(water_dist, water_2k)) + geom_abline() + geom_point(alpha = 0.5)

```

Makes sense that water distance is essentially equal - we've just added
2km radius and many values are tens of km away.

## NPEV detection by site and month

### Overall

```{r}

es |> 
  ggplot(aes(year, fill = !is.na(npev))) + 
  geom_bar() + 
  labs(x = NULL, y = "No. Samples", fill = "NPEV+")

es |> 
  group_by(year) |>
  summarise(n = n(),
            npev_pos = sum(!is.na(npev)),
            lo = Hmisc::binconf(npev_pos, n)[2],
            hi = Hmisc::binconf(npev_pos, n)[3]) |>
  ungroup() |> 
  ggplot(aes(year, npev_pos/n, ymin = lo, ymax = hi)) + 
  geom_errorbar(width = 0.2) +
  geom_point() +
  ylim(c(0,1)) +
  labs(fill = "NPEV+", 
       x = NULL, y = "Proportion positive", 
       title = "Sample NPEV positivity by year",
       subtitle = "95% binomial confidence intervals")

es |> 
  group_by(month) |>
  summarise(n = n(),
            npev_pos = sum(!is.na(npev)),
            lo = Hmisc::binconf(npev_pos, n)[2],
            hi = Hmisc::binconf(npev_pos, n)[3]) |>
  ungroup() |> 
  ggplot(aes(month, npev_pos/n, ymin = lo, ymax = hi, col = (npev_pos/n > 0.5))) + 
  geom_errorbar(width = 0.2) +
  geom_point() +
  geom_vline(xintercept = ymd(paste0(2021:2025,"-01-01")), 
             lty = "dashed", col = "grey") +
  ylim(c(0,1)) +
  labs(fill = "NPEV+", 
       x = NULL, y = "Proportion positive",
       col = "NPEV+ > 50%",
       title = "Sample NPEV positivity by month",
       subtitle = "95% binomial confidence intervals")


es |> 
  mutate(site_no = as.numeric(as.factor(site_id))) |> 
  group_by(admin_1, site_id, site_no) |>
  summarise(n = n(),
            npev_pos = sum(!is.na(npev)),
            lo = Hmisc::binconf(npev_pos, n)[2],
            hi = Hmisc::binconf(npev_pos, n)[3]) |> 
  group_by(admin_1) |>
  mutate(adm1_br = ifelse(row_number() == 1, site_no - 0.5, NA)) |> 
  ungroup() |>
  ggplot(aes(site_no, npev_pos/n, ymin = lo, ymax = hi, col = (npev_pos/n > 0.5))) + 
  geom_errorbar(width = 0.2) +
  geom_point() +
  geom_vline(aes(xintercept = adm1_br)) +
  ylim(c(0,1)) +
  labs(fill = "NPEV+", 
       x = NULL, y = "Proportion positive",
       col = "NPEV+ > 50%",
       title = "Sample NPEV positivity by site",
       subtitle = "95% binomial confidence intervals")

```

```{r}

es |> 
  group_by(site_id, x, y) |> 
  summarise(p_npev = sum(!is.na(npev))/n(),
            pop = mean(catchment_pop_5k)) |> 
  ungroup() |> 
  st_as_sf(coords = c("x","y"), crs = 4326) -> tmp
ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = tmp, aes(col = p_npev*100, size = pop), pch = 15, alpha = 0.6) + 
  scale_colour_viridis_c(limits = c(0,100)) +
  # theme(legend.position = c(0.9,0.25)) + 
  labs(fill = NULL,
       colour = "% NPEV+",
       size = "Estimated\ncatchment",
       title = "Percentage of NPEV+ samples per site",
       caption = "At least one sample collected Jan 2021-Dec 2024.") +
  theme_void()

```

### By time

```{r}

es |> 
  ggplot(aes(month, fill = !is.na(npev))) + 
  geom_bar() + 
  labs(x = "Collection month", y = "No. samples", fill = "NPEV+")

es |> 
  mutate(collection_month = month,
         month = lubridate::month(collection_month, label = T)) |> 
  group_by(site_id, x, y, year, month, collection_month) |> 
  summarise(n_samples = n(),
            p_npev = sum(!is.na(npev))/n_samples) |> 
  ungroup() |> 
  st_as_sf(coords = c("x","y"), crs = 4326) -> tmp

tmp |> 
  mutate(year = as.factor(year)) |> 
  ggplot(aes(month, p_npev, group = year)) + 
  geom_line(alpha = 0.5, col = "grey") + 
  geom_smooth(aes(col = year), method = "loess") +
  labs(x = NULL, y = "Prop. NPEV+", col = "Year")

```

### By site and time

```{r}

es |> 
  group_by(site_id, x, y, year) |> 
  summarise(n_samples = n(),
            p_npev = sum(!is.na(npev))/n_samples) |> 
  ungroup() |> 
  st_as_sf(coords = c("x","y"), crs = 4326) -> tmp

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326), fill = "white") + 
  geom_sf(data = st_jitter(tmp, factor = 0.0005), 
          aes(col = p_npev, alpha = n_samples)) + 
  facet_wrap(~year) + 
  scale_color_viridis_c(trans = "sqrt", direction = 1) +  
  labs(caption = "Points jittered for visibility.",
       title = "NPEV positivity per site, per year",
       alpha = "No. samples",
       colour = "Proportion") + 
  theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank())

ggplot() +
  geom_sf(data = shape0 |> st_transform(4326)) +
  geom_sf(data = st_jitter(tmp, factor = 0.0005), aes(fill = (p_npev >= 0.5)), pch = 22) +
  facet_wrap(~year) +
  theme_minimal() +
  labs(caption = "Points jittered for visibility.",
       fill = NULL,
       title = "At least 50% NPEV positivity") + 
  theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank())

```

# Modelling

## Make analysis dataset

To begin with, look at just total per year given site characteristics
and location.

Aggregate samples to site/month/year totals and define spatial geometry:

```{r}

es |> 
  mutate(collection_month = month,
         month = lubridate::month(month, label = T)) |> 
  group_by(collection_month, month, year, site_id, x, y) |> 
  summarise(
    n_samples = n(),
    n_cvdpv2 = sum(final_class == "cVDPV2", na.rm = T),
    n_npev = sum(!is.na(npev))) |> 
  ungroup()|> 
  mutate(p_npev = n_npev/n_samples,
         p_cvdpv2 = n_cvdpv2/n_samples) |> 
  left_join(es_sites_utm |> st_drop_geometry() |> select(site_id, catchment_2k)) -> es_agg_mth

es_pt_mth <- es_agg_mth |> 
  st_as_sf(coords = c("x","y"), crs = 4326) 

es_pt_mth |> 
  group_by(year, site_id, catchment_2k) |> 
  summarise(across(n_samples:n_npev, sum)) |> 
  ungroup()|> 
  mutate(p_npev = n_npev/n_samples,
         p_cvdpv2 = n_cvdpv2/n_samples) -> es_pt_yr

```

```{r}

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_jitter(es_pt_yr, factor = 0.0005), pch = 22, fill = "white") + 
  facet_wrap(~year) +  
  theme_minimal() + 
  labs(caption = "Points jittered for visibility.",
       title = "Unique sites reporting samples between Jan 2021 and Oct 2024") 

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_jitter(es_pt_yr, factor = 0.0005), aes(col = p_npev), cex = 0.8) + 
  facet_wrap(~year) + 
  scale_color_viridis_c(trans = "sqrt") + 
  theme_minimal() + 
  labs(caption = "Points jittered for visibility.",
       title = "NPEV positivity per site, per year",
       colour = "Proportion") 

ggplot() +
  geom_sf(data = shape0 |> st_transform(4326)) +
  geom_sf(data = st_jitter(es_pt_yr, factor = 0.0005), aes(fill = (p_npev >= 0.5)), pch = 22) +
  facet_wrap(~year) +
  theme_minimal() +
  labs(caption = "Points jittered for visibility.",
       fill = NULL,
       title = "At least 50% NPEV positivity")

```

Initially few sites with \>50% NPEV positivity rate per year, but
increasing over time.

Visualise these predictors against empirical logit of p[NPEV]:

```{r}

es_pt_yr |> 
  mutate(elogitp = log((n_npev+0.5)/(n_samples-n_npev+0.5))) |> 
  bind_cols(st_coordinates(es_pt_yr)) |> 
  st_drop_geometry() |> 
  select(catchment_2k, year,
         X, Y,
         elogitp) %>% 
  pivot_longer(-elogitp) %>% 
  ggplot(aes(value, elogitp)) + 
  geom_jitter(alpha = 0.5) + 
  geom_smooth() + 
  facet_wrap(~name, scales = "free_x") + 
  theme_bw()

```

Also for monthly positivity:

```{r}

es_pt_mth |> 
  mutate(elogitp = log((n_npev+0.5)/(n_samples-n_npev+0.5)),
         month = as.numeric(month)) |> 
  bind_cols(st_coordinates(es_pt_mth)) |>
  st_drop_geometry() |> 
  select(month, year, 
         catchment_2k,
         X, Y,
         elogitp) |>
  pivot_longer(-elogitp) |>
  ggplot(aes(value, elogitp)) + 
  geom_jitter(alpha = 0.2) + 
  geom_smooth() + 
  facet_wrap(~name, scales = "free_x") + 
  theme_bw()

```

Distinct bump in NPEV-positivity during early summer.

## Monthly

```{r}

plot_es_npev_pos <- function(yr){
 ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_jitter(filter(es_pt_mth, year == yr), factor = 0.0005), 
          aes(col = p_npev, size = n_samples), alpha = 0.5) + 
  facet_wrap(~month) + 
  scale_color_viridis_c(trans = "sqrt") + 
  theme_minimal() + 
  theme(axis.text = element_blank()) +  
  labs(caption = "Points jittered for visibility.",
       title = paste0("NPEV positivity per site, per month (",yr,")"),
       colour = "Proportion\npositive",
       alpha = "No.\nsamples") -> p
  
  return(p)
}

lapply(c(2021:2024), plot_es_npev_pos)

```

Look at seasonal trend by year:

```{r}

es_pt_mth |> 
  mutate(year = as.factor(year)) |> 
  ggplot(aes(month, p_npev, group = year)) + 
  geom_line(alpha = 0.5, col = "grey") + 
  geom_smooth(aes(col = year), method = "loess")

```

Between 2021-23 there was an overall increase in NPEV positivity, with a
similar late-summer peak each year. The increase towards the peak is
more gradual than the decrease after. In 2021 and 2023 we see a small
bump coming up to July but this is not observed in 2022, which appears
more wobbly.

The 2024 trend is quite different to the other years, generally
decreasing with a small bump in June.

```{r}

plot_es_npev_pos <- function(yr){
 ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_jitter(filter(es_sf, year == yr), factor = 0.0005), 
          aes(col = p_npev, size = n_samples), alpha = 0.5) + 
  facet_wrap(~month) + 
  scale_color_viridis_c(trans = "sqrt") + 
  theme_minimal() + 
  theme(axis.text = element_blank()) +  
  labs(caption = "Points jittered for visibility.",
       title = paste0("NPEV positivity per site, per month (",yr,")"),
       colour = "Proportion\npositive",
       alpha = "No.\nsamples") -> p
  
  return(p)
}

lapply(c(2021:2024), plot_es_npev_pos)

```

Try fitting a model by month:

```{r}

modeldata_m <- bind_cols(es_pt_mth, st_coordinates(es_pt_mth)) |> 
  mutate(m = as.numeric(month),
         year_f = as.factor(year)) |> 
  group_by(site_id) |> 
  mutate(sid = cur_group_id() |>  as.factor()) |> 
  ungroup()

m4 <- gam(data = modeldata_m,
          weights = n_samples,
          p_npev ~ s(catchment_2k, k = 3) + # s(built_2k, k = 3) + s(water_dist, k = 3) + 
            s(m, by = year_f, k = 5) + year_f + 
            s(X, Y, bs = "gp"),
          family = binomial())

summary(m4)
plot(m4, pages = 1)

```

Want to look at effect of recent rainfall on npev detection, and in
particular when in time this effect was significant e.g.
here[<https://r.qcbs.ca/workshop08/book-en/changing-the-basis-function.html>]
and
here[<https://aslopubs.onlinelibrary.wiley.com/doi/10.4319/lo.2009.54.6_part_2.2529>].

<br>

# Incorporate NPEV in AFP notifications

```{r}

afp |> 
  group_by(guid,x,y) |> 
  count(name = "n_unique") |> 
  ungroup() |> 
  group_by(n_unique) |> 
  count()

afp |> 
  select(guid, x, y) |> 
  unique() -> afp_locs

table(afp$npev)
```

```{r}

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_as_sf(afp_locs, coords = c("x","y"), crs = 4326), pch = 22, fill = "white") + 
  labs(title = "Given locations of notified AFP cases",
       caption = "Jan 2021-Dec 2024.")

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_as_sf(afp |> filter(!is.na(npev)), coords = c("x","y"), crs = 4326), 
          aes(fill = onset_date), pch = 22, alpha = 0.4) + 
  scale_fill_viridis_c() + 
  theme(legend.position = c(0.9,0.25)) + 
  labs(fill = NULL,
       title = "NPEV-positive notified AFP cases",
       caption = "Jan 2021-Dec 2024.")
```

Why do black borders of points seem to recreate district boundaries?

# Modelling

First fit a model for NPEV positivity given these covariates:

```{r}

# Exclude obs with missing built extent
modeldata <- bind_cols(es_pt_yr, st_coordinates(es_pt_yr)) |> 
  group_by(site_id) |> 
  mutate(sid = cur_group_id() |>  as.factor()) |> 
  ungroup()

m1 <- gam(data = modeldata, 
          weights = n_samples,
          p_npev ~ s(catchment_2k, k = 3) + year,
          family = binomial())

summary(m1)
plot(m1, pages = 1)

```

Then look at residual spatial correlation:

```{r}

resid_dev <- resid(m1, "deviance")
plot(resid_dev, main = "Deviance Residuals", ylab = "Residuals", xlab = "Index")

modeldata |> 
  select(site_id, geometry) |> 
  mutate(resid_m1 = resid_dev) |> 
  as_Spatial() -> m1_resids_sp

# Compute the empirical variogram
vgm_residuals <- variogram(resid_m1 ~ 1, data = m1_resids_sp)
plot(vgm_residuals, main = "Empirical Variogram of Residuals", xlab = "Distance", ylab = "Semivariance")

```

Some evidence of residual spatial correlation. Also an undulating
pattern with increasing distance? Could this be down to non-spatial
heterogeneity between sites?

```{r}

m2 <- gam(data = modeldata,
          weights = n_samples,
          p_npev ~ s(catchment_2k, k = 3) + year + s(sid, bs = "re"),
          family = binomial())

summary(m2)
plot(m2, pages = 1)

AIC(m1, m2)

# Look at spatial correlation in site random effects
m2_re <- gratia::smooth_estimates(m2) |> filter(.smooth == "s(sid)")
plot(m2_re$sid, m2_re$.estimate, main = "Fitted site-level random effect", ylab = "Estimate", xlab = "Site")

modeldata |> 
  select(sid, site_id, geometry) |> 
  full_join(select(m2_re, sid, .estimate, .se)) |> 
  as_Spatial() -> m2_re_sp

# Compute the empirical variogram
vgm_re <- variogram(.estimate ~ 1, data = m2_re_sp)
plot(vgm_re, main = "Empirical Variogram of site-level random effects", xlab = "Distance", ylab = "Semivariance")

vgm_re_se <- variogram(.se ~ 1, data = m2_re_sp)
plot(vgm_re_se, main = "Empirical Variogram of site-level random effects (Std Err)", xlab = "Distance", ylab = "Semivariance")

```

Apparently stronger spatial trend in site-level random effects - also
catchment and built area have lost significance (not surprising).

```{r}

m3 <- gam(data = modeldata,
          weights = n_samples,
          p_npev ~ s(catchment_2k, k = 3) + year + s(X, Y, bs = "gp") + s(sid, bs = "re"),
          family = binomial())

summary(m3)
plot(m3, pages = 1)

AIC(m1, m2, m3)

# Plot the 2D smooth term effect (it won't be a full 2D plot but marginal effects)
plot(m3, select = 4)

m3_re <- gratia::smooth_estimates(m3) |> filter(.smooth == "s(sid)")
plot(m3_re$sid, m3_re$.estimate, main = "Fitted site-level random effect", ylab = "Estimate", xlab = "Site")

modeldata |> 
  select(sid, site_id, geometry) |> 
  full_join(select(m3_re, sid, .estimate, .se)) |> 
  as_Spatial() -> m3_re_sp

vgm_re_3 <- variogram(.estimate ~ 1, data = m3_re_sp)
plot(vgm_re_3, main = "Empirical Variogram of site-level random effects", xlab = "Distance", ylab = "Semivariance")
```

Spatial smooth doesn't fit to anything now that site-level REs included.

```{r}

augment(m1, 
        type.predict = "link",
        type.residuals = "pearson") %>%
  {bind_cols(modeldata[,'geometry'],.)} %>%
  ggplot(data= .) +
  geom_sf(aes(col = .resid), pch = 19, alpha = 0.7) +
  scale_colour_gradient2(name = "Residuals", limits = c(-4,4)) +
  theme_bw() 
augment(m2, 
        type.predict = "link",
        type.residuals = "pearson") %>%
  {bind_cols(modeldata[,'geometry'],.)} %>%
  ggplot(data= .) +
  geom_sf(aes(col = .resid), pch = 19, alpha = 0.7) +
  scale_colour_gradient2(name = "Residuals", limits = c(-4,4)) +
  theme_bw() 

```

### Plot prediction over space

Just look at variation captured in the GP first - i.e. set other
covariates to average:

```{r}

# Create a grid of points for prediction
bbox <- st_bbox(shape0 |> st_transform(4326))
x_range <- seq(bbox[1], bbox[3], length.out = 100)
y_range <- seq(bbox[2], bbox[4], length.out = 100)
grid <- expand.grid(X = x_range, Y = y_range)

# Set other covariates (z and w) to their mean values
mean_catch <- mean(modeldata$catchment_2k)
mean_built <- mean(modeldata$built_2k)
mean_water <- mean(modeldata$water_dist)

# Add the constant covariates to the grid and year
grid$catchment_2k <- mean_catch
grid$built_2k <- mean_built
grid$water_dist <- mean_water
grid$year <- 2024

grid$z_pred <- predict(m2, newdata = grid)

# Plot the smooth term effect using a viridis color scale
ggplot() +
  geom_tile(data = grid, aes(x = X, y = Y, fill = z_pred)) +
  geom_sf(data = st_union(shape2) |> st_transform(4326), color = "black", alpha = 0) +
  scale_fill_viridis_c() + 
  labs(title = "2D Smooth Term Effect from GAM", x = "X Coordinate", y = "Y Coordinate") +
  theme_minimal()

```
