---
title: "ES sites in Nigeria - 2021-24"
output: html_notebook
---

# Background

Confidence in the interruption of poliovirus transmission is highly dependent on assumed sensitivity of environmental surveillance (ES), in particular in the case of cVDPV2 (low rate of clinical disease). Previous work has explored how the strength of evidence against ongoing circulation is influenced by variation in which sites (in which locations, relative to the distribution of population) are reporting samples at what time. This work also incorporated detection rate of NPEVs as another source of variation between sites, hence variation in the strength of evidence for absence in that location when no poliovirus has been detected. 

The latter indicator was considered only with respect to a fixed target rate defined by the GPEI (>50%), which is used to monitor site performance and inform their opening/closure. This means that sites detecting NPEVs in an average of 50% of collected samples over the previous 12 months are considered as performing at their "maximum", with our measurement of sensitivity scaled down when this value falls below 50%. In fact, the 50% threshold is arbitrarily defined and doesn't necessarily align with the expected prevalence of NPEVs in the population, in different locations and at different times of year. 

In this setting, the nature of collection sites varies considerably between natural waterways, informal drainage channels, and more formal sewage networks. These different types of site will likely have very different population catchment and virus detection sensitivity. Sites are, however, only defined by name and GPS location. In order to distinguish different site types, we can look at the land cover surrounding the site and the proximity/intersection with natural water bodies. 

Ultimately, we want to infer the relationship between probability of NPEV detection and:
 - Population size within Xkm
 - Dominant land cover type within Xkm
 - Proximity to natural waterways/bodies
 - Elevation within Xkm compared to elevation at site? 
[- Proportion of "flooded vegetation" within Xkm]

For the last point, we are interested in the influence of seasonal flooding on NPEV detection. This may alter the usual catchment of a site, as well as diluting/contaminating samples with run-off from other regions. However, this is not a significant consideration in the context of Nigeria and should instead be explored in a more flood-prone setting such as Pakistan.

The WHO establishes ES sites to supposedly capture ~100,000 people, however in reality this often isn't the case. Previous work has demonstrated significant inconsistency in the interpretation of site catchment. Different sites will capture different-sized areas according to the nature of the site, therefore it is not clear what catchment radius is appropriate to assumed for analysis. This could be investigated by 

Methodological approach: 
 - Fit GAM of P(NPEV) ~ catch_pop + lc_type + proximity_water 
 - Investigate anisotropy in the correlation structure via directional variograms


## Set up environment and load data

```{r}
pacman::p_load(tidyverse, sf, janitor, terra, exactextractr, mgcv, gstat, broom)
theme_set(theme_minimal())

# Inputs
indir <- "../polio-spec-models/inputs/Nigeria - manuscript/nga/"

# Outputs
outdir <- "data"
if(!dir.exists(outdir)) dir.create(outdir, recursive = T)

proj_local <- "EPSG:32631"

# ES linelist
es <- readRDS(file.path(indir,"es.rds"))

# AFP linelist
afp <- readRDS(file.path(indir,"afp.rds"))

# Population raster
poprast  <- rast("data/nga_ppp_2020_UNadj.tif") |> 
  terra::project(proj_local)

# Administrative boundaries
shape2 <- readRDS(file.path(indir, "shape2.rds")) |> 
  st_transform(proj_local)

shape0 <- st_union(shape2)

```

## Describe unique sites

```{r}

es |> 
  select(guid, admin_0_vizname, admin_1_vizname, admin_2_vizname, 
         site_id:site_comment, 
         x, y) |> 
  unique() -> es_sites

glimpse(es_sites)

```
In total, `r nrow(es_sites)` unique sites active at some point between `r min(es$collection_date)` and `r max(es$collection_date)`.

Check uniqueness of coordinates per site code:
```{r}

nrow(es_sites)
n_distinct(es_sites$site_code)

```

```{r}

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_as_sf(es_sites, coords = c("x","y"), crs = 4326), pch = 22, fill = "white") + 
  labs(title = "Wastewater sampling sites in Nigeria",
       caption = "At least one sample collected Jan 2021-Dec 2024.")

es |> 
  group_by(site_code, x, y) |> 
  summarise(p_npev = mean(!is.na(npev))) |> 
  ungroup() -> tmp 
ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_as_sf(tmp, coords = c("x","y"), crs = 4326), aes(fill = p_npev), pch = 22) + 
  scale_fill_viridis_c(limits = c(0,1)) +
  theme(legend.position = c(0.9,0.25)) + 
  labs(fill = NULL,
       title = "Overall proportion NPEV+ samples per site",
       caption = "At least one sample collected Jan 2021-Dec 2024.")

```
## Site characteristics

Apparently `r sum(es_sites$site_status == "CLOSED")` sites classified as "closed", however at least one sample must have been collected 2021-24 to appear in this dataset. Which are these samples?

```{r}

janitor::tabyl(es_sites$site_status)

```

```{r}

es_sites |> 
  filter(site_status == "CLOSED") -> closed_sites

print(closed_sites)

es |> 
  filter(site_code %in% closed_sites$site_code) |> 
  select(site_code, site_name, site_status, site_comment, sample_condition, npev, final_class) |> 
  print()

```

All are commented with "poor enterovirus isolation", however all samples from this site marked as in "good" condition with decent proportion NPEV+. Perhaps now closed after these samples were taken?
-> Decide to retain supposedly "closed" sites.

### Site names

Can we glean anything from site names? 
- Some in particular refer to bridges, specific streams/rivers, community hubs (e.g. school/mosque/camp/police station) and road intersections (junction/roundabout)
- There is one formal wastewater treatment plant

What proportion of samples from identifiable site types are NPEV-positive?

```{r}

es |> 
  mutate(site_name = tolower(site_name),
         canal = grepl("canal",site_name),
         bridge = grepl("bridge",site_name),
         str_river = grepl("stream",site_name) | grepl("river",site_name),
         comm_site = grepl("school",site_name) | grepl("market",site_name) | grepl("mosque",site_name),
         camp = grepl("camp", site_name),
         station = grepl("station",site_name), 
         junction = grepl("junction",site_name),
         wwtp = grepl("treatment plant",site_name),
         no_type = !(canal|bridge|str_river|comm_site|camp|station|junction|wwtp)) -> tmp

tmp |> summarise(across(canal:no_type, sum)) 

tmp |> 
  pivot_longer(canal:no_type, names_to = "type") |> 
  filter(value) |> 
  ggplot(aes(type, fill = !is.na(npev))) + 
  geom_bar() + 
  labs(fill = "NPEV+", x = NULL, y = "Count", title = "Sample count and positivity by rough site type")

```

Not really helpful... 

### Catchment population size

```{r}

es_sites_utm <- st_as_sf(es_sites, coords = c("x","y"), crs = 4326) |> 
  st_transform(proj_local)

es_buff <- st_buffer(es_sites_utm, dist = units::set_units(2, "km"))

# Population within 2km of each site:
es_sites_utm$catchment_2k <- exactextractr::exact_extract(poprast, es_buff, "sum") 

# Total population per admin2 area:
shape2$total_pop <- exactextractr::exact_extract(poprast, shape2, "sum")
    
hist(es_sites_utm$catchment_2k)
summary(es_sites_utm$catchment_2k)

```
On average, `r mean(es_sites_utm$catchment_2k, na.rm = T)` people reside within 2km of an individual collection site, with a total of `r sum(es_sites_utm$catchment_2k)` within 2km of *any* site (~`r round(sum(es_sites_utm$catchment_2k)/sum(shape2$total_pop)*100)`\% of the national population).  

```{r}

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = es_sites_utm |> st_transform(4326), aes(size = catchment_2k, colour = catchment_2k), pch = 15, alpha = 0.7) + 
  scale_colour_viridis_c(trans = "sqrt", labels = scales::label_comma()) +
  theme(legend.position = c(0.9,0.25)) + 
  guides(size = "none") + 
  labs(colour = NULL,
       title = "Catchment population within 2km of site")

```

### Land cover within catchment area

Load rasters of land use and distance to waterways:
[Note that these from WorldPop are quite out of date - should be replaced with ESRI land cover estimates]

```{r}

water_dist  <- rast("data/nga_osm_dst_waterway_100m_2016.tif") |> 
  terra::project(proj_local)

built_extent  <- rast("data/nga_bsgme_v0a_100m_2020.tif") |> 
  terra::project(proj_local)

```

```{r}

plot(water_dist)
plot(built_extent)

```
- Built extent raster is binary, but seems to be aggregated to a different resolution in autoplot hence non-binary values. 
- Water distance raster is in km to the nearest *major* waterway.

First look at the point value at each site (create a small buffer to still use exact_extract):

```{r}

es_buff_sm <- st_buffer(es_sites_utm, dist = units::set_units(50, "m"))

es_sites_utm$built_extent <- exactextractr::exact_extract(built_extent, es_buff_sm, "mean")
es_sites_utm$water_dist <- exactextractr::exact_extract(water_dist, es_buff_sm, "mean") 

summary(es_sites_utm$built_extent)
hist(es_sites_utm$built_extent)

summary(es_sites_utm$water_dist)
hist(es_sites_utm$water_dist)

```
Also look at average value within 2km catchment: 

```{r}

es_sites_utm$built_2k <- exactextractr::exact_extract(built_extent, 
                                                      es_buff,
                                                      "mean") 
es_sites_utm$water_2k <- exactextractr::exact_extract(water_dist, 
                                                      es_buff,
                                                      "mean") 

ggplot(es_sites_utm, aes(built_extent, built_2k)) + geom_abline() + geom_point(alpha = 0.5) 
ggplot(es_sites_utm, aes(water_dist, water_2k)) + geom_abline() + geom_point(alpha = 0.5)
```
Makes sense that water distance is essentially equal - we've just added 2km radius and many values are tens of km away. 

## Make analysis dataset

To begin with, look at just total per year given site characteristics and location. 

Aggregate samples to site/month/year totals and define spatial geometry: 

```{r}

es |> 
  mutate(collection_month = month,
         month = lubridate::month(month, label = T)) |> 
  group_by(collection_month, month, year, site_id, x, y) |> 
  summarise(
    n_samples = n(),
    n_cvdpv2 = sum(final_class == "cVDPV2", na.rm = T),
    n_npev = sum(!is.na(npev))) |> 
  ungroup()|> 
  mutate(p_npev = n_npev/n_samples,
         p_cvdpv2 = n_cvdpv2/n_samples) |> 
  left_join(es_sites_utm |> st_drop_geometry() |> select(site_id, catchment_2k:water_2k)) -> es_agg_mth

es_pt_mth <- es_agg_mth |> 
  st_as_sf(coords = c("x","y"), crs = 4326) 

es_pt_mth |> 
  group_by(year, site_id, catchment_2k, built_extent, water_dist, built_2k, water_2k) |> 
  summarise(across(n_samples:n_npev, sum)) |> 
  ungroup()|> 
  mutate(p_npev = n_npev/n_samples,
         p_cvdpv2 = n_cvdpv2/n_samples) -> es_pt_yr

```


```{r}

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_jitter(es_pt_yr, factor = 0.0005), pch = 22, fill = "white") + 
  facet_wrap(~year) +  
  theme_minimal() + 
  labs(caption = "Points jittered for visibility.",
       title = "Unique sites reporting samples between Jan 2021 and Oct 2024") 

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_jitter(es_pt_yr, factor = 0.0005), aes(col = p_npev), cex = 0.8) + 
  facet_wrap(~year) + 
  scale_color_viridis_c(trans = "sqrt") + 
  theme_minimal() + 
  labs(caption = "Points jittered for visibility.",
       title = "NPEV positivity per site, per year",
       colour = "Proportion") 

ggplot() +
  geom_sf(data = shape0 |> st_transform(4326)) +
  geom_sf(data = st_jitter(es_pt_yr, factor = 0.0005), aes(fill = (p_npev >= 0.5)), pch = 22) +
  facet_wrap(~year) +
  theme_minimal() +
  labs(caption = "Points jittered for visibility.",
       fill = NULL,
       title = "At least 50% NPEV positivity")

```
Initially few sites with >50% NPEV positivity rate per year, but increasing over time.  

Visualise these predictors against empirical logit of p[NPEV]:

```{r}

es_pt_yr |> 
  mutate(elogitp = log((n_npev+0.5)/(n_samples-n_npev+0.5))) |> 
  bind_cols(st_coordinates(es_pt_yr)) |> 
  st_drop_geometry() |> 
  select(catchment_2k:water_2k, year,
         X, Y,
         elogitp) %>% 
  pivot_longer(-elogitp) %>% 
  ggplot(aes(value, elogitp)) + 
  geom_jitter(alpha = 0.5) + 
  geom_smooth() + 
  facet_wrap(~name, scales = "free_x") + 
  theme_bw()

```

Also for monthly positivity: 

```{r}

es_pt_mth |> 
  mutate(elogitp = log((n_npev+0.5)/(n_samples-n_npev+0.5)),
         month = as.numeric(month)) |> 
  bind_cols(st_coordinates(es_pt_mth)) |>
  st_drop_geometry() |> 
  select(month, year, 
         catchment_2k:water_2k, 
         X, Y,
         elogitp) |>
  pivot_longer(-elogitp) |>
  ggplot(aes(value, elogitp)) + 
  geom_jitter(alpha = 0.2) + 
  geom_smooth() + 
  facet_wrap(~name, scales = "free_x") + 
  theme_bw()

```

Distinct bump in NPEV-positivity during early summer.

### How many samples are collected per site?

Look at sampling frequency per site:

```{r}

tabyl(es_pt_mth$n_samples)

```
The majority of sites are reporting one sample per month, with the rest reporting two. On very few occasions has a site reported more than two samples in a given month. Which sites are these? Do some consistently report more than others?

```{r}

tabyl(st_drop_geometry(es_pt_mth), month, n_samples)

ggplot(st_drop_geometry(es_pt_mth)) + 
  geom_bar(aes(month, fill = as.factor(n_samples))) + 
  labs(y = "Number of sites", x = "Month", fill = "Sample\nfrequency")

```
Fewer sites sampling twice-monthly during summer.

# Modelling

First fit a model for NPEV positivity given these covariates:

```{r}

# Exclude obs with missing built extent
modeldata <- bind_cols(es_pt_yr, st_coordinates(es_pt_yr)) |> 
  filter(!is.na(built_extent)) |> 
  group_by(site_id) |> 
  mutate(sid = cur_group_id() |>  as.factor()) |> 
  ungroup()

m1 <- gam(data = modeldata, 
          weights = n_samples,
          p_npev ~ s(catchment_2k, k = 3) + s(built_2k, k = 3) + s(water_dist, k = 3) + year,
          family = binomial())

summary(m1)
plot(m1, pages = 1)

```
Then look at residual spatial correlation: 

```{r}

resid_dev <- resid(m1, "deviance")
plot(resid_dev, main = "Deviance Residuals", ylab = "Residuals", xlab = "Index")

modeldata |> 
  select(site_id, geometry) |> 
  mutate(resid_m1 = resid_dev) |> 
  as_Spatial() -> m1_resids_sp

# Compute the empirical variogram
vgm_residuals <- variogram(resid_m1 ~ 1, data = m1_resids_sp)
plot(vgm_residuals, main = "Empirical Variogram of Residuals", xlab = "Distance", ylab = "Semivariance")

```
Doesn't seem to be any evidence of residual spatial correlation. However, there does seem to be an undulating pattern with increasing distance? Could this be down to non-spatial heterogeneity between sites?


```{r}

m2 <- gam(data = modeldata,
          weights = n_samples,
          p_npev ~ s(catchment_2k, k = 3) + s(built_2k, k = 3) + s(water_dist, k = 3) + year + s(sid, bs = "re"),
          family = binomial())

summary(m2)
plot(m2, pages = 1)

AIC(m1, m2)

# Look at spatial correlation in site random effects
m2_re <- gratia::smooth_estimates(m2) |> filter(.smooth == "s(sid)")
plot(m2_re$sid, m2_re$.estimate, main = "Fitted site-level random effect", ylab = "Estimate", xlab = "Site")

modeldata |> 
  select(sid, site_id, geometry) |> 
  full_join(select(m2_re, sid, .estimate, .se)) |> 
  as_Spatial() -> m2_re_sp

# Compute the empirical variogram
vgm_re <- variogram(.estimate ~ 1, data = m2_re_sp)
plot(vgm_re, main = "Empirical Variogram of site-level random effects", xlab = "Distance", ylab = "Semivariance")

vgm_re_se <- variogram(.se ~ 1, data = m2_re_sp)
plot(vgm_re_se, main = "Empirical Variogram of site-level random effects (Std Err)", xlab = "Distance", ylab = "Semivariance")

```
Apparently stronger spatial trend in site-level random effects - also catchment and built area have lost significance (not surprising).

```{r}

m3 <- gam(data = modeldata,
          weights = n_samples,
          p_npev ~ s(catchment_2k, k = 3) + s(built_2k, k = 3) + s(water_dist, k = 3) + year + s(X, Y, bs = "gp") + s(sid, bs = "re"),
          family = binomial())

summary(m3)
plot(m3, pages = 1)

AIC(m1, m2, m3)

# Plot the 2D smooth term effect (it won't be a full 2D plot but marginal effects)
plot(m3, select = 4)

m3_re <- gratia::smooth_estimates(m3) |> filter(.smooth == "s(sid)")
plot(m3_re$sid, m3_re$.estimate, main = "Fitted site-level random effect", ylab = "Estimate", xlab = "Site")

modeldata |> 
  select(sid, site_id, geometry) |> 
  full_join(select(m3_re, sid, .estimate, .se)) |> 
  as_Spatial() -> m3_re_sp

vgm_re_3 <- variogram(.estimate ~ 1, data = m3_re_sp)
plot(vgm_re_3, main = "Empirical Variogram of site-level random effects", xlab = "Distance", ylab = "Semivariance")
```

Spatial smooth doesn't fit to anything now that site-level REs included. 

```{r}

augment(m1, 
        type.predict = "link",
        type.residuals = "pearson") %>%
  {bind_cols(modeldata[,'geometry'],.)} %>%
  ggplot(data= .) +
  geom_sf(aes(col = .resid), pch = 19, alpha = 0.7) +
  scale_colour_gradient2(name = "Residuals", limits = c(-4,4)) +
  theme_bw() 
augment(m2, 
        type.predict = "link",
        type.residuals = "pearson") %>%
  {bind_cols(modeldata[,'geometry'],.)} %>%
  ggplot(data= .) +
  geom_sf(aes(col = .resid), pch = 19, alpha = 0.7) +
  scale_colour_gradient2(name = "Residuals", limits = c(-4,4)) +
  theme_bw() 

```

### Plot prediction over space

Just look at variation captured in the GP first - i.e. set other covariates to average: 

```{r}

# Create a grid of points for prediction
bbox <- st_bbox(shape0 |> st_transform(4326))
x_range <- seq(bbox[1], bbox[3], length.out = 100)
y_range <- seq(bbox[2], bbox[4], length.out = 100)
grid <- expand.grid(X = x_range, Y = y_range)

# Set other covariates (z and w) to their mean values
mean_catch <- mean(modeldata$catchment_2k)
mean_built <- mean(modeldata$built_2k)
mean_water <- mean(modeldata$water_dist)

# Add the constant covariates to the grid and year
grid$catchment_2k <- mean_catch
grid$built_2k <- mean_built
grid$water_dist <- mean_water
grid$year <- 2024

grid$z_pred <- predict(m2, newdata = grid)

# Plot the smooth term effect using a viridis color scale
ggplot() +
  geom_tile(data = grid, aes(x = X, y = Y, fill = z_pred)) +
  geom_sf(data = st_union(shape2) |> st_transform(4326), color = "black", alpha = 0) +
  scale_fill_viridis_c() + 
  labs(title = "2D Smooth Term Effect from GAM", x = "X Coordinate", y = "Y Coordinate") +
  theme_minimal()

```

## 2022 only

```{r}

plot_es_npev_pos <- function(yr){
 ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_jitter(filter(es_pt_mth, year == yr), factor = 0.0005), 
          aes(col = p_npev, alpha = n_samples), cex = 0.8) + 
  facet_wrap(~month) + 
  scale_color_viridis_c(trans = "sqrt") + 
  theme_minimal() + 
  theme(axis.text = element_blank()) +  
  labs(caption = "Points jittered for visibility.",
       title = paste0("NPEV positivity per site, per month (",yr,")"),
       colour = "Proportion\npositive",
       alpha = "No.\nsamples") -> p
  
  return(p)
}

lapply(c(2021:2024), plot_es_npev_pos)

```
Look at seasonal trend by year:

```{r}

es_pt_mth |> 
  mutate(year = as.factor(year)) |> 
  ggplot(aes(month, p_npev, group = year, col = year)) + 
  geom_line() + 
  geom_smooth()

```
Try fitting a model by month:

```{r}

modeldata_m <- bind_cols(es_pt_mth, st_coordinates(es_pt_mth)) |> 
  filter(!is.na(built_extent)) |> 
  mutate(m = as.numeric(month)) |> 
  group_by(site_id) |> 
  mutate(sid = cur_group_id() |>  as.factor()) |> 
  ungroup()
m4 <- gam(data = modeldata_m,
          weights = n_samples,
          p_npev ~ s(catchment_2k, k = 3) + # s(built_2k, k = 3) + s(water_dist, k = 3) + 
            s(m, by = year, bs = "cc") + year + 
            s(X, Y, bs = "gp"),
          family = binomial())

summary(m4)
plot(m4, pages = 1)

```
Want to look at effect of recent rainfall on npev detection, and in particular when in time this effect was significant e.g. here[https://r.qcbs.ca/workshop08/book-en/changing-the-basis-function.html] and here[https://aslopubs.onlinelibrary.wiley.com/doi/10.4319/lo.2009.54.6_part_2.2529].

<br>

# Incorporate NPEV in AFP notifications

```{r}

afp |> 
  group_by(guid,x,y) |> 
  count(name = "n_unique") |> 
  ungroup() |> 
  group_by(n_unique) |> 
  count()

afp |> 
  select(guid, x, y) |> 
  unique() -> afp_locs

table(afp$npev)
```

A few duplicated locations - these could be hospitals/facilities but doesn't seem to be a massive problem? With 42k unique locations can we assume these are households/communities?

```{r}

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_as_sf(afp_locs, coords = c("x","y"), crs = 4326), pch = 22, fill = "white") + 
  labs(title = "Given locations of notified AFP cases",
       caption = "Jan 2021-Dec 2024.")

ggplot() + 
  geom_sf(data = shape0 |> st_transform(4326)) + 
  geom_sf(data = st_as_sf(afp |> filter(!is.na(npev)), coords = c("x","y"), crs = 4326), aes(fill = onset_date), pch = 22) + 
  scale_fill_viridis_c() + 
  theme(legend.position = c(0.9,0.25)) + 
  labs(fill = NULL,
       title = "NPEV-positive notified AFP cases",
       caption = "Jan 2021-Dec 2024.")
```