---
title: 'Part 1: District NPEV prevalence model'
output:
  html_document:
    df_print: paged
---

```{r}

pacman::p_load(tidyverse, here, gtsummary, sf, spdep, brms, tidybayes, units, loo)
theme_set(theme_minimal())

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

dir <- "data/Pakistan"

source("./utils/fit_npafp_model.R")
source("./utils/check_model_fit.R")

# NP-AFP analysis data
data <- readRDS(here(dir, "afp_analysis.rds"))

# Shapefiles and adjacency matrix
shape2 <- readRDS(here(dir,"shape2.rds")) 
W <- readRDS(here(dir,"W.rds"))

```

# Descriptive

```{r}

data |> 
  tbl_summary(include = c(n_afp, n_npafp, n_npev, npafp_r, npev_npafp_p),
              by = year,
              label = list(n_afp = "Total AFP notifications", 
                           n_npafp = "Non-polio AFP notifications", 
                           n_npev = "NPEV positives", 
                           npafp_r = "Non-polio AFP rate (per 100,000 U15s)", 
                           npev_npafp_p = "Crude NPEV prevalence")) |> 
  add_overall() |> 
  as_gt()

```

Plot crude NPEV prevalence among notified non-polio AFP cases, per district and over time:

```{r}

# Annual NPEV+ rate by district
data |> 
  group_by(guid, year) |> 
  summarise(n_npafp = sum(n_npafp),
            n_npev = sum(n_npev),
            r_npev = n_npev/n_npafp) |>
  ungroup() |>
  right_join(shape2) |> 
  ggplot(aes(fill = r_npev, geometry = SHAPE)) + 
  geom_sf(col = NA) +
  facet_wrap(~year) +
  scale_fill_viridis_c(name = "NPEV+\nPrevalence", trans = "sqrt") + 
  theme_void() -> p_obs_dist_yr

# Seasonal trend, by year
data |> 
  group_by(month_of_year, year) |> 
  summarise(n_npafp = sum(n_npafp),
            n_npev = sum(n_npev),
            r_npev = n_npev/n_npafp) |>
  ungroup() |>
  ggplot(aes(x = month_of_year, y = r_npev, 
             group = factor(year), col = factor(year), fill = factor(year))) + 
  geom_point() + 
  geom_smooth() + 
  # guides(fill = "none") +
  labs(x = NULL, y = "NPEV+ Prevalence", col = "Year", fill = "Year") -> p_obs_mth_yr

# Seasonal trend, by province
data |> 
  group_by(month_of_year, adm1_name) |> 
  summarise(n_npafp = sum(n_npafp),
            n_npev = sum(n_npev),
            r_npev = n_npev/n_npafp) |>
  ungroup() |>
  ggplot(aes(x = month_of_year, y = r_npev, 
             group = adm1_name, col = adm1_name, fill = adm1_name)) + 
  # geom_jitter(alpha = 0.2) + 
  geom_smooth(alpha = 0.5) + 
  labs(x = NULL, y = "NPEV+ Prevalence") -> p_obs_mth_prov

# Seasonal trend, by district
data |> 
  group_by(month_of_year, guid) |> 
  summarise(n_npafp = sum(n_npafp),
            n_npev = sum(n_npev),
            r_npev = n_npev/n_npafp) |>
  ungroup() |>
  ggplot(aes(x = month_of_year, y = r_npev, 
             group = guid)) + 
  # geom_jitter(alpha = 0.2) + 
  geom_smooth(alpha = 0.1, lwd = 0.2, se = F) + 
  labs(x = NULL, y = "NPEV+ Prevalence") -> p_obs_mth_dist

p_obs_mth_yr
p_obs_dist_yr
p_obs_mth_prov
p_obs_mth_dist

```

Overall decrease year on year, with more pronounced seasonality in 2021. 

# Modelling 

## Model setup

1. NPEV prevalence model

Number of NPEV+ AFP cases in district $i$ at time $t$:

$$ Y_{i,t} \sim  Binomial(M_{i,t}, p_{i,t}) $$
where $M_{i,t}$ is the number of notified AFP cases and $p_{i,t}$ is the estimated prevalence of NPEV in this sampled population.

The prevalence per district-month is dependent on population density, spatial correlation between neighbouring districts, a seasonal trend and a 1st order random walk:

$$ logit(p_{i,t}) = \beta_0 + \beta_1*[popdens] + f^{BYM}(i) + f^{cc}(month(t)) + f^{RW1}(t)$$
The predicted prevalence and SE for each district-month will be extracted for input as a predictor of NPEV detection probability in subsequent ES model. 

2. ES detection model

Binary detection of NPEV in sample $j$ from site $s$ at time $t$:

$$ X_{j,s,t} \sim  Bernoulli(q_{j,s,t}) $$

where $q_{j,s,t}$ is the probability of detecting NPEV, given district prevalence $p_{i,t}$.

$$ logit(q_{j,s,t}) = \gamma_0 + \gamma_s*log(p_{i,t}) + f_{s,t}(month(t)) $$
The relation $i~j$ will be defined according to three options: 
- Site $j$ point location is within district $i$
- Majority watershed of site $j$ intersects with district $i$
- Any watershed of site $j$ intersects with district $i$ where $i \in I$, using a weighted average of $p_{i,t}$ over all $i \in I$. [Question: How to average SE?]

Other notes:
Based on Kroiss et al. analysis, NPEV prevalence was around 25% per month. 
 => N(-1,1) informative prior on the global intercept. 

## Model Fitting

Setting up data for fitting:

```{r}

fitdata <- mutate(data, 
                  month_of_year = as.numeric(month_of_year),
                  year_F = factor(year),
                  guid_F = factor(guid, levels = unique(data$guid)),
                  t_F = factor(t, levels = 1:max(data$t)),
                  # standardise pop density for stability
                  log_pop_dens = log(guid_pop_dens),
                  log_pop_dens_s = (log_pop_dens - mean(log_pop_dens))/sd(log_pop_dens)) 

```

Custom RW prior:
```{r}

prior_rw <- function(idx_inter = 1, 
                     sigma_inter, 
                     prior_scale_sigma_rw,...
                     ) {
  
  pr <- prior_string(paste0("rw1(", idx_inter, ", ", sigma_inter, ", sigma_rw)"), ...) +
    prior_string(paste0("target += normal_lpdf(sigma_rw| 0, ", prior_scale_sigma_rw, ")"), check=FALSE)
  
  sv_rw1 <- stanvar(scode = "real rw1_lpdf(vector alpha, int idx_inter, real sigma_inter, real sigma_rw) {
    int N = num_elements(alpha);
    return normal_lpdf(alpha[2:N] - alpha[1:N - 1] | 0, sigma_rw) + normal_lpdf(alpha[idx_inter]| 0, sigma_inter);
  }", 
                    block = "functions")
  
  sv_sigma_rw <- stanvar(scode="real<lower=0> sigma_rw;", 
                         block="parameters")
  
  list(prior=pr, stanvar=sv_rw1 + sv_sigma_rw)
}

```
### Baseline model {.tabset}

Linear effect of log population density (standardised), periodic spline, and temporal RW.

#### Formula and priors

```{r}

f_base = bf(
  n_npev | trials(n_npafp) ~ lin + rw,
  lin ~ 0 + log_pop_dens_s +
    # Periodic spline on 1-12 for seasonality
    s(month_of_year, bs = "cc", k = 12),
  rw ~ 1 + t_F, nl=TRUE, center=FALSE)

# Check default priors given this formula
get_prior(f_base, fitdata, family = binomial("logit"))

# Define the RW prior
rw_prior <- prior_rw(1, 5, 0.5, class="b", nlpar="rw")

```

#### Fitting

```{r}

fit_base <- brm(f_base, data = fitdata,
           family = "binomial",
               prior = rw_prior$prior,
               stanvars = rw_prior$stanvar,
               refresh=250,
               empty=FALSE, init=0.2, 
           adapt_delta = 0.95) 

fit_base
```

#### Check fit
```{r}

# Prior predictive checks
plot_pp_checks(fit_base)

# Trace plots
plot(fit_base)

# Residuals
check_residuals(fit_base, 
                data = fitdata, 
                shapefile = shape2)
```

### Adding a spatial component {.tabset}

Test spatial fit without RW (just seasonality and IID reff on t):

#### Formula and priors
```{r}

f_bym = bf(
  n_npev | trials(n_npafp) ~ 0 + log_pop_dens_s +
    car(W, gr = guid_F, type = "bym2") + 
    s(month_of_year, bs = "cc", k = 12) + (1|t)
  )

# Check default priors given this formula
get_prior(f_bym, 
          data = fitdata, data2 = list(W = W), 
          family = binomial("logit"))

```

#### Fitting
```{r}

fit_bym <- brm(f_bym, 
               data = fitdata,
               data2 = list(W = W),
               family = "binomial",
               refresh=250,
               empty=FALSE, init=0.2, 
           adapt_delta = 0.95) 

fit_bym

```

#### Check fit
```{r}

plot_pp_checks(fit_bym)
plot(fit_bym)

```

### Combining BYM with RW {.tabset}

Full BYM-RW model.

#### Formula and priors
```{r}

f_bym_rw = bf(
  n_npev | trials(n_npafp) ~ lin + rw,
  lin ~ 0 + log_pop_dens_s +
    s(month_of_year, bs = "cc", k = 12) + 
    car(W, gr = guid_F, type = "bym2"),
  rw ~ 1 + t_F, nl=TRUE, center=FALSE)

# Check default priors given this formula
get_prior(f_bym_rw, 
          data = fitdata, data2 = list(W = W), 
          family = binomial("logit"))

```

#### Fitting
```{r}

f_bym_rw <- brm(f_bym_rw, 
               data = fitdata,
               data2 = list(W = W),
               family = "binomial",
               prior = rw_prior$prior,
               stanvars = rw_prior$stanvar,
               refresh=250,
               empty=FALSE, init=0.2, 
           adapt_delta = 0.95) 

f_bym_rw

```

#### Check fit
```{r}

plot_pp_checks(f_bym_rw)
plot(f_bym_rw)

```

### Model comparison
```{r}

fits <- list(fit_base, fit_bym, fit_bym_rw)
saveRDS(fits, "output/fits.rds")

lapply(fits, 
       add_criterion, 
       c("waic","loo"), moment_match = T)

# Compare models
loo::loo_compare(fits)

```

Select final model for prediction:

```{r}

fit_prev <- fits[[3]]

```

## Extract posterior draws and plot 

```{r}

posterior_samples <- as_draws_df(fit_prev)  

# Extract RW component 
rw_t <- posterior_samples |>  select(starts_with("r_t"))  

# Summarise across draws
rw_summary <- rw_t |> 
  pivot_longer(cols = everything(), names_to = "time", values_to = "rw_value") |>
  mutate(time = as.numeric(gsub("r_t\\[", "", gsub(",Intercept\\]", "", time)))) |>
  group_by(time) |> 
  summarise(
    mean_rw = mean(rw_value),
    lower = quantile(rw_value, 0.025),
    upper = quantile(rw_value, 0.975)
  ) 
  # # Transform to probability scale
  # mutate(across(mean_rw:upper, \(x) exp(x)/(1+exp(x))))

ggplot(rw_summary, aes(x = time, y = mean_rw)) +
  geom_line(color = "steelblue") +
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              alpha = 0.2, fill = "steelblue") +
  theme_minimal() +
  labs(
    title = "Estimated first-order random walk",
    x = "Time",
    y = "Estimate"
  )

```

-> Downward shift at the end of 2022? 

# Predict prevalence by district-month

Set at least one NP-AFP case per district-month to force prediction > 0.

```{r}

newdata <- fitdata
newdata$n_npafp[newdata$n_npafp == 0] <- 1

# Extract predictions
pred_prev <- posterior_epred(fit_prev, 
                             newdata = newdata, 
                             scale = "linear")

# Summarise 
pred_summary <- as.data.frame(t(apply(pred_prev, 
                                      2, 
                                      quantile, 
                                      probs = c(0.025, 0.5, 0.975))))
colnames(pred_summary) <- c("lower", "median", "upper")

# Add to data and scale to prevalence
data_wpred <- newdata |> 
  mutate(pred_lo = pred_summary$lower,
         pred_med = pred_summary$median,
         pred_hi = pred_summary$upper, 
         pred_mean = apply(pred_prev, 2, mean),
         pred_sd = apply(pred_prev, 2, sd),
         across(starts_with("pred"), \(x) x/n_npafp)) 

```

## Check 

```{r}

data_wpred |> 
  select(pred_lo:pred_sd) |> 
  summary()

summary(data_wpred$pred_med - data_wpred$pred_mean)

```

## Plotting

### By month and district

```{r}

data_wpred |> 
  ggplot(aes(month, group = guid)) + 
  geom_jitter(aes(y = npev_npafp_p), 
             col = "grey", alpha = 0.5, cex = 0.2) +
  geom_ribbon(aes(ymin = pred_lo, ymax = pred_hi), 
              alpha = 0.2, fill = "lightsteelblue") + 
  geom_line(aes(y = pred_med), alpha = 0.2, col = "steelblue4") + 
  # geom_line(aes(y = pred_mean), col = "red", alpha = 0.1) +
  labs(y = "Estimated prevalence", x = "Month", title= "Predicted median & quantiles")

data_wpred |> 
  ggplot(aes(month, pred_mean, 
             ymin = pred_mean-2*pred_sd, 
             ymax = pred_mean+2*pred_sd, group = guid)) + 
  geom_jitter(aes(y = npev_npafp_p), 
             col = "grey", alpha = 0.5, cex = 0.2) +
  geom_ribbon(alpha = 0.2, fill = "lightsteelblue") + 
  geom_line(alpha = 0.2, col = "steelblue4") + 
  labs(y = "Estimated prevalence", x = "Month", title = "Predicted mean & SD")

```
Minimal difference between median and mean => use mean & SD as inputs to ES detection model. 

### Annual averages

```{r}

data_wpred |> 
  group_by(guid, year) |> 
  summarise(pred_mean = mean(pred_mean),
            pred_sd = mean(pred_sd)) |>
  ungroup() |>
  right_join(shape2) -> tmp

tmp |> 
  ggplot(aes(fill = pred_mean, geometry = SHAPE)) + 
  geom_sf(col = NA) +
  facet_wrap(~year) +
  scale_fill_viridis_c(name = "Predicted NPEV+\nPrevalence") + 
  theme_void()
tmp |> 
  ggplot(aes(fill = pred_sd, geometry = SHAPE)) + 
  geom_sf(col = NA) +
  facet_wrap(~year) +
  scale_fill_viridis_c(name = "Predicted NPEV+\nPrevalence - SD") + 
  theme_void()

```
### Monthly

Look at spatial trends per quarter, across years.

```{r}

data_wpred |> 
  filter(period %% 0.25 == 0) |> 
  right_join(shape2) -> tmp

tmp |> 
  ggplot(aes(fill = pred_mean, geometry = SHAPE)) + 
  geom_sf(col = NA) +
  facet_wrap(~month) +
  scale_fill_viridis_c(name = "Predicted NPEV+\nPrevalence") + 
  theme_void()
tmp |> 
  ggplot(aes(fill = pred_sd, geometry = SHAPE)) + 
  geom_sf(col = NA) +
  facet_wrap(~month) +
  scale_fill_viridis_c(name = "Predicted NPEV+\nPrevalence - SD") + 
  theme_void()

```
