---
title: 'Part 1: District NPEV prevalence'
subtitle: '1B: Modelling'
output:
  html_document:
    df_print: paged
---

```{r}

# remotes::install_github("Pakillo/DHARMa.helpers")

pacman::p_load(tidyverse, here, gtsummary, sf, spdep, rstan, brms, tidybayes, bayesplot, units, loo, posterior, DHARMa, DHARMa.helpers)
theme_set(theme_minimal())

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

dir <- "data/Pakistan/analysis/prevalence"
outdir <- "output/prevalence/exploratory"

source("../utils/check_model_fit.R")
source("../utils/plot_re_fcns.R")

# NP-AFP analysis data
data <- readRDS(here(dir, "afp_analysis.rds"))

# Shapefiles 
shape2 <- readRDS(here(dir,"../shape2.rds")) |> st_transform(4326)

```

```{r}
plot_pp_checks <- function(fit){
  
  check_plots <- list(pp_check(fit, type = "bars"), 
                      pp_check(fit, type = "scatter_avg"),
                      pp_check(fit, "rootogram"))  
  return(check_plots)
}
```

## Model specification

1. NPEV prevalence model

Use a beta-binomial model as some sample counts are small. 

Number of NPEV+ AFP cases in district $i$ at time $t$:

$$ Y_{i,t} \sim  Beta-Binomial(M_{i,t}, p_{i,t}) $$
where $M_{i,t}$ is the number of notified non-polio AFP cases and $p_{i,t}$ is the estimated prevalence of NPEV in this sampled population.

The prevalence per district-month is assumed dependent on a seasonal trend, spatial correlation between neighbouring districts, and other spatio-temporal noise:

$$ logit(p_{i,t}) = \beta_0 + f^{cc}(month(t)) + f^{BYM}(i) + f^{IID}(t) + f^{IID}(i:t)$$
This model will be constructed incrementally to justify the addition of complexity from a baseline seasonal trend + noise model. The objective is to explain with this model the majority of the variability in observed prevalence per month:district. Our purpose is not to predict out-of-sample but simply to obtain smoothed estimates 

The predicted prevalence and SE for each district-month will be extracted for input as a predictor of NPEV detection probability in subsequent ES model. 

Other notes:
Based on Kroiss et al. analysis, NPEV prevalence was around 25% per month. 
 => N(-1,1) informative prior on the global intercept. 

## Data setup 

Setting up data for fitting and prediction:
- preddata including all district:month combinations for later prediction
- fitdata including only those with >0 npafp (as prevalence is undefined)

```{r}

# Set up data
preddata <- mutate(data, 
                  # Scale time variables
                  time = t,
                  t = (time - min(time))/diff(range(time)),
                  m = as.numeric(month_of_year),
                  month_of_year = (m - min(m))/diff(range(m)),
                  # standardise pop density for stability
                  log_pop_dens = log(guid_pop_dens) |> as.numeric(),
                  log_pop_dens_s = (log_pop_dens - mean(log_pop_dens))/sd(log_pop_dens)) |> 
  rename(guid_old = guid) |> 
  left_join(shape2 |> st_drop_geometry() |> select(guid_old, guid))

# Exclude zero trial rows for fitting (will still predict for these later) 
fitdata <- filter(preddata, n_npafp > 0)
write_rds(preddata, here(dir, "preddata.rds"))
write_rds(fitdata, here(dir, "fitdata.rds"))

# Define adm1-guid key for later matching:
prov_guid <- preddata |> 
  select(adm1_name, guid) |> 
  distinct()

```

## Modelling

### Baseline {.tabset}

Overall intercept, IID district and month random effects, and periodic spline for seasonal trend.

#### Formula and priors

```{r}

f_base = bf(
  n_npev | trials(n_npafp) ~ 
    # Overall average prevalence
    1 + 
    # Random deviations per province, district and time
    (1|adm1_name) + (1|guid) + (1|t) +
    # Periodic spline for seasonality
    s(month_of_year, bs = "cc", k = 12))

# Check default priors given this formula
get_prior(f_base, fitdata, family = binomial("logit"))

```

"Flat" prior is improper, essentially infinitely small density across -Inf to +Inf. 

```{r}

rexp(10000,2) |> hist()
rexp(10000,1) |> hist()
boot::inv.logit(rstudent_t(10000, 3, 0, 2.5)) |> hist()
boot::inv.logit(rnorm(10000,-1,1)) |> hist()

# prior_base <- c(prior(normal(-1,1), class = Intercept), # More informative on logit scale
#                 prior(exponential(1), class = sd, coef = "Intercept", group = "adm1_name"), # Conservative on sd of reffs, tighter for district than province
#                 prior(exponential(1.5), class = sd, coef = "Intercept", group = "guid"),
#                 prior(exponential(1), class = sd, coef = "Intercept", group = "t"),
#                 prior(exponential(1), class = sds))

prior_base <- c(prior(normal(-1,1), class = Intercept), # More informative on logit scale
                prior(exponential(2), class = sd),
                prior(exponential(2), class = sds))

```

#### Prior checks

```{r}

pp_base <- brm(f_base, 
               data = fitdata,
               data2 = list(W = W),
               family = "binomial",
               prior = prior_base,
               sample_prior = "only",
               file = here(outdir, "prior_base.rds")) 

```

```{r}

summary(pp_base)

plot_pp_checks(pp_base)

```

#### Fitting

```{r}

fit_base <- brm(f_base, 
               data = fitdata,
               data2 = list(W = W),
               family = "binomial",
               prior = prior_base,
               refresh = 250,
               empty = FALSE, init = 0.2,
               iter = 3000, chains = 4, 
               cores = parallel::detectCores(),
               control = list(adapt_delta = 0.96),
               save_pars = save_pars(all = T),
               file_refit = "on_change",
               file = here(outdir, "fit_base.rds")) 

```

#### Check fit

```{r}
summary(fit_base)
bayes_R2(fit_base)
```

```{r}

mcmc_areas(fit_base, 
           pars = get_variables(fit_base)[1:5])

# Prior predictive checks
plot_pp_checks(fit_base)

# Trace plots
plot(fit_base)

# Residuals
# check_residuals(fit_base, 
#                 data = fitdata, 
#                 shapefile = shape2)

check_residuals(fit_base)

```

Fitted random effects:

```{r}

plot_re_adm1(fit_base)
plot_re_guid(fit_base)
plot_re_adm1_guid(fit_base, prov_guid, prov_lines = F)

plot_re_t(fit_base)

```

```{r}

plot(conditional_smooths(fit_base))

```

### Adding spatial correlation {.tabset}

#### Formula and priors
```{r}

f_bym = bf(
  n_npev | trials(n_npafp) ~ 
    # Overall average prevalence
    1 + 
    # Random deviations by province and time
    (1|adm1_name) + (1|t) +
    # Correlated/uncorrelated random deviations by district 
    car(W, gr = guid, type = "bym2") + 
    # Seasonal temporal spline
    s(month_of_year, bs = "cc", k = 12) 
  )

# Check default priors given this formula
get_prior(f_bym, 
          data = fitdata, data2 = list(W = W), 
          family = binomial("logit"))

rbeta(10000,1,1) |> hist()
rbeta(10000,2,5) |> hist()
rbeta(10000,0.5,0.5) |> hist()

# PC prior for sdcar s.t. P((1/sqrt(tau)) > 1) = 0.01 => exp(theta)
prior_bym <- c(prior_base,
               # Tighten prior on province with added structure in district term? (nearby districts more similar which counters province-level differences)
               # prior(exponential(), coef = sd, group = "adm1_name"),
               # Prior for rho that favours uncorrelated error
               prior(beta(1,2), class = rhocar),
               # Tighter prior on CAR sd than adm1 random intercept
               prior(exponential(2), class = sdcar))

```

#### Fitting
```{r}

fit_bym <- update(fit_base, 
                  formula = f_bym,
                  prior = prior_bym,
                  newdata = fitdata,
                  data2 = list(W = W),
                  recompile = T,
                  control = list(adapt_delta = 0.97),
                  file_refit = "on_change",
                  file = here(outdir,"fit_bym.rds")) 

```

#### Check fit
```{r}
summary(fit_bym)
bayes_R2(fit_bym)
```

```{r}

vars <- get_variables(fit_bym)
mcmc_areas(fit_bym, 
           pars = get_variables(fit_bym)[1:5])

plot_pp_checks(fit_bym)
plot(fit_bym)

# check_residuals(fit_bym, 
#                 data = fitdata, 
#                 shapefile = shape2)
check_residuals(fit_bym)

```

Explore temporal AC a bit further:
```{r}

# In IID reff:
fit_bym |> 
    gather_draws(r_t[t,]) |> 
    median_qi() |> 
    left_join(select(fitdata, t, month) |> distinct()) -> tmp1

# In residuals
tmp2 <- mutate(fitdata, resid = residuals(fit_bym, type = "pearson")[,1])

tmp2 |> 
  group_by(guid) |> 
  summarise(acf1 = acf(resid, plot = FALSE)$acf[2],
            acf2 = acf(resid, plot = FALSE)$acf[3],
            n = n()) -> acf_guid

  
acf(ts(tmp1$.value))
pacf(ts(tmp1$.value))

summary(acf_guid$acf1)
hist(acf_guid$acf1, main = "Lag-1 autocorrelation by district", xlab = "ACF(1)")

```
- ACF of temporal reff indicates auto-correlation up to lag ~6m -> long-term smooth trend.   
- Average ACF of residuals is approx zero -> roughly independent over time within each district. 

```{r}

plot_re_adm1(fit_bym)

fit_bym %>%
  gather_draws(rcar[guid]) %>%
  median_qi() |> 
  left_join(prov_guid) |> 
  mutate(component = "Total") -> summ_car

summ_car |> 
  arrange(adm1_name, guid) |> 
  mutate(did = row_number()) |> 
  ggplot(aes(did, .value,
             ymin = .lower, ymax = .upper,
             col = adm1_name)) + 
  geom_hline(yintercept = 0, col = "grey") +
  geom_errorbar() + 
  geom_point() +
  labs(x = "District", y = "Posterior estimate", col = "Province",
       subtitle = "Total BYM component") +
  scale_colour_viridis_d(option = "turbo") +
  theme(axis.text.x = element_blank())

fit_bym %>%
  gather_draws(nszcar[guid]) %>%
  median_qi() |> 
  left_join(prov_guid) |> 
  mutate(component = "Non-spatial") -> summ_car_ns

summ_car |> 
  bind_rows(summ_car_ns) |> 
  arrange(adm1_name, guid) |> 
  mutate(did = row_number()) |> 
  ggplot(aes(did, .value,
             ymin = .lower, ymax = .upper,
             col = adm1_name,
             alpha = component)) + 
  geom_hline(yintercept = 0, col = "grey") +
  geom_errorbar() + 
  geom_point() +
  labs(x = "District", y = "Posterior estimate", 
       col = "Province", alpha = "Component",
       subtitle = "Conditional auto-regressive (CAR) effect") +
  scale_colour_viridis_d(option = "turbo") +
  theme(axis.text.x = element_blank())

fit_bym %>%
  spread_draws(r_adm1_name[adm1_name, Intercept],rcar[guid]) %>% 
  left_join(mutate(prov_guid,guid = row_number()), by = "guid") |> 
  filter(adm1_name.x == adm1_name.y) |> 
  median_qi(r_guid = r_adm1_name + rcar) -> tmp

tmp |> 
  left_join(mutate(prov_guid,guid = row_number())) |> 
  arrange(adm1_name, guid) |> 
  mutate(did = row_number()) |> 
  ggplot(aes(did, r_guid,
             ymin = .lower, ymax = .upper,
             col = adm1_name)) + 
  geom_hline(yintercept = 0, col = "grey") +
  geom_errorbar() + 
  geom_point() +
  labs(x = "District", y = "Posterior estimate", 
       col = "Province",
       subtitle = "Total district deviation from average") +
  scale_colour_viridis_d(option = "turbo") +
  theme(axis.text.x = element_blank())

```
Lots of uncertainty in the non-spatial (IID) CAR component. 

```{r}

plot(conditional_smooths(fit_bym)) 

plot_re_t(fit_bym)

```

### Adding temporal correlation (long-term spline) {.tabset}

In addition to seasonal trend, in place of purely IID temporal random effects since these show temporal auto-correlation.

#### AR(1) term
```{r}

f_bym_ar = bf(
  n_npev | trials(n_npafp) ~ 
    # Overall average prevalence
    1 + 
    # Random deviations by province
    (1|adm1_name) + (1|t) +
    # Correlated/uncorrelated random deviations by district 
    car(W, gr = guid, type = "bym2") + 
    # AR1
    ar(t, gr = guid, p = 1) +
    # Seasonal temporal spline
    s(month_of_year, bs = "cc", k = 12) 
  )

# Check default priors given this formula
get_prior(f_bym_ar, 
          data = fitdata, data2 = list(W = W), 
          family = binomial("logit"))

# Add weakly-informative uniform prior on ar parameter
prior_bym_ar <- c(prior_bym, 
                  prior(uniform(-1,1), class = ar))

```

```{r}

fit_bym_ar <- update(fit_bym, 
                  formula = f_bym_ar,
                  prior = prior_bym_ar,
                  newdata = fitdata,
                  data2 = list(W = W),
                  recompile = T,
                  control = list(adapt_delta = 0.96),
                  file_refit = "on_change",
                  file = here(outdir, "fit_bym_ar.rds"))

```

```{r}
summary(fit_bym_ar)
bayes_R2(fit_bym_ar)
check_residuals(fit_bym_ar)
```

Plot temporal components:

```{r}

plot(conditional_smooths(fit_bym_ar)) 
plot_re_t(fit_bym_ar)

```

#### Non-cyclic spline

Long-term smooth temporal trend:

```{r}

f_bym_st = bf(
  n_npev | trials(n_npafp) ~ 
    # Overall average prevalence
    1 + 
    # Random deviations by province and time
    (1|adm1_name) + (1|t) +
    # Correlated/uncorrelated random deviations by district 
    car(W, gr = guid, type = "bym2") + 
    # Temporal spline
    s(t, k = 10) +
    # Seasonal temporal spline
    s(month_of_year, bs = "cc", k = 12) 
  )

# Check default priors given this formula
get_prior(f_bym_st, 
          data = fitdata, data2 = list(W = W), 
          family = binomial("logit"))

# Tighten prior on IID to avoid competing with other temporal effects
prior_bym_st <- c(prior_bym, 
                  prior(exponential(2.5), class = sd, coef = "Intercept", group = "t"))

```

```{r}

fit_bym_st <- update(fit_bym, 
                  formula = f_bym_st,
                  prior = prior_bym_st,
                  newdata = fitdata,
                  data2 = list(W = W),
                  recompile = T,
                  control = list(adapt_delta = 0.98),
                  file_refit = "on_change",
                  file = here(outdir, "fit_bym_st.rds")) 

```

```{r}
summary(fit_bym_st)
bayes_R2(fit_bym_st)
```

```{r}

mcmc_areas(fit_bym_st, 
           pars = get_variables(fit_bym_st)[1:8])

plot_pp_checks(fit_bym_st)
plot(fit_bym_st)

# check_residuals(fit_bym_st, 
#                 data = fitdata, 
#                 shapefile = shape2)

check_residuals(fit_bym_st)
check_residuals(fit_bym_st, var = fitdata$t)
check_residuals(fit_bym_st, var = fitdata$adm1_name)

```

Temporal components:

```{r}

plot(conditional_smooths(fit_bym_st)) 
plot_re_t(fit_bym_st)

```
Not better than IID wrt R2 and temporal spline seems poorly identified? But does separate long term and seasonal trends from residual variability. IID effects still seems to show auto-correlation. 

Look at the residuals over time by district:

```{r}

# In IID reff:
fit_bym_st |> 
    gather_draws(r_t[t,]) |> 
    median_qi() |> 
    left_join(select(fitdata, t, month) |> distinct()) -> tmp1

# In residuals
tmp2 <- mutate(fitdata,
               resid = residuals(fit_bym_st, type = "pearson")[,1])

tmp2 |> 
  group_by(guid) |> 
  summarise(acf1 = acf(resid, plot = FALSE)$acf[2],
            acf2 = acf(resid, plot = FALSE)$acf[3],
            n = n()) -> acf_guid

acf(ts(tmp1$.value))
pacf(ts(tmp1$.value))

summary(acf_guid$acf1)
hist(acf_guid$acf1, main = "Lag-1 autocorrelation by district", xlab = "ACF(1)")

```

Autocorrelation seems to have been accounted for.

### Non-cyclic spline by guid {.tabset}

#### Formula and priors
```{r}

f_bym_st_guid = bf(
  n_npev | trials(n_npafp) ~     
    # Overall average prevalence
    1 + 
    (1|adm1_name) + (1|t) +
    # Long-term spline over time
    s(t, bs = "tp", k = 10) +
    # Deviation by guid (penalise wiggliness now we're fitting many)
    s(t, guid, bs = "fs", k = 6) + 
    # Correlated/uncorrelated random deviations by district 
    car(W, gr = guid, type = "bym2") + 
    # Seasonal temporal splines overall and by province
    s(month_of_year, bs = "cc", k = 12) 
  )

# Check default priors given this formula
get_prior(f_bym_st_guid, 
          data = fitdata, data2 = list(W = W), 
          family = binomial("logit"))

prior_bym_st_guid <- c(prior_bym_st,
                       prior(normal(0,1), class = "b"),
                       prior(exponential(2), class = sds, coef = 's(t, bs = "tp", k = 10)'),
                       # Tighter prior on spline sd for district deviations
                       prior(exponential(2.5), class = sds, coef = 's(t, guid, bs = "fs", k = 6)'))

```

#### Prior check

```{r}

pp_bym_st_guid <- brm(f_bym_st_guid, 
               data = fitdata,
               data2 = list(W = W),
               family = "binomial",
               prior = prior_bym_st_guid,
               sample_prior = "only",
               file = here(outdir, "prior_bym_st_guid.rds")) 

```

```{r}

summary(pp_bym_st_guid)

plot_pp_checks(pp_bym_st_guid)

```

#### Fitting

Increase iterations here due to more complex model:

```{r}

fit_bym_st_guid <- update(fit_bym_st, 
                  formula = f_bym_st_guid,
                  prior = prior_bym_st_guid,
                  newdata = fitdata,
                  data2 = list(W = W),
                  cores = parallel::detectCores(),
                  iter = 4000, warmup = 2000,
                  save_pars = save_pars(all = TRUE),
                  control = list(adapt_delta = 0.98),
                  recompile = T,
                  file = here(outdir, "fit_bym_st_guid.rds"))

```

#### Check fit
```{r}
summary(fit_bym_st_guid)
bayes_R2(fit_bym_st_guid)
```

```{r}

mcmc_areas(fit_bym_st_guid, 
           pars = get_variables(fit_bym_st_guid)[1:12])

plot_pp_checks(fit_bym_st_guid)

# check_residuals(fit_bym_st_guid, 
#                 data = fitdata, 
#                 shapefile = shape2)

check_residuals(fit_bym_st_guid)
check_residuals(fit_bym_st_guid, var = fitdata$t)
check_residuals(fit_bym_st_guid, var = fitdata$adm1_name)

```

Setup temporal plots:
```{r}

fit_bym_st_guid |> 
    conditional_smooths() |>
    plot(plot = F) -> plots
  
  ggplot_build(plots[[1]]) -> p1
  ggplot_build(plots[[2]]) -> p2
  ggplot_build(plots[[3]]) -> p3

p2$plot$data |> 
  left_join(select(fitdata, guid, guid_old) |> distinct()) |> 
  rename(District = guid_old) |> 
  ggplot(mapping = p2$plot$mapping) + 
  # geom_hline(yintercept = 0, lty = "dashed", col = "darkgrey") +
  geom_ribbon(aes(group = District), alpha = 0.01, fill = "steelblue", col = NA) +
  geom_line(aes(group = District), alpha = 0.2, col = "steelblue4") +
  guides(col = "none",fill = "none") +
  labs(x = NULL, y = "District-specific trend") -> p2_comb

p2$plot$labels <- list(x = NULL, y = "Overall seasonal trend")
p2$plot$scales$scales[[1]] <- scale_x_discrete(limits = month.abb) 
p2$plot$theme <- theme(axis.text.x = element_text(angle = 45, hjust = 1))

fit_bym_st_guid %>%
  gather_draws(r_t[t, Intercept]) %>%
  median_qi() |> 
  # Need to redefine integer time as rounding error in scaled t seems to be messing up join
  mutate(time = row_number()) |> 
  left_join(select(fitdata, time, month) |> distinct()) -> summ_r_t

```

```{r}

p1
p2_comb
p3

summ_r_t |>  
  ggplot(aes(month, .value,
             ymin = .lower, ymax = .upper)) + 
  geom_hline(yintercept = 0, col = "grey") +
  geom_errorbar() + 
  geom_point() +
  labs(x = "Month", y = "Posterior estimate", subtitle = "Temporal random effect")
  
```

### Non-cyclic spline by province {.tabset}

#### Formula and priors
```{r}

# f_bym_st_prov = bf(
#   n_npev | trials(n_npafp) ~     
#     # Overall average prevalence
#     1 + 
#     # Random deviations by province and time
#     (1|adm1_name) + (1|t) + 
#     # Long-term spline over time
#     s(t, bs = "tp", k = 10) +
#     # Deviation by guid (penalise wiggliness now we're fitting many)
#     s(t, adm1_name, bs = "fs", k = 6) + 
#     # Correlated/uncorrelated random deviations by district 
#     car(W, gr = guid, type = "bym2") + 
#     # Seasonal temporal splines by province
#     s(month_of_year, bs = "cc", k = 12) 
#   )

```

#### Fitting

```{r}

# fit_bym_st_prov <- update(fit_bym_st, 
#                   formula = f_bym_st_prov,
#                   # Use same priors as before
#                   prior = prior_bym_st_guid,
#                   newdata = fitdata,
#                   data2 = list(W = W),
#                   cores = parallel::detectCores(),
#                   iter = 4000,
#                   control = list(adapt_delta = 0.97),
#                   recompile = T,
#                   file = here(outdir,"fit_bym_st_prov.rds")) 

```

#### Check fit
```{r}
# summary(fit_bym_st_prov)
# bayes_R2(fit_bym_st_prov)
```

```{r}


# fit_bym_st_prov |> 
#     conditional_smooths() |>
#     plot(plot = F) -> plots
#   
#   ggplot_build(plots[[1]]) -> p1
#   ggplot_build(plots[[2]]) -> p2
# 
# p1$plot$data |> 
#   rename(Province = adm1_name) |> 
#   ggplot(mapping = p1$plot$mapping) + 
#   # geom_hline(yintercept = 0, lty = "dashed", col = "darkgrey") +
#   geom_ribbon(aes(group = Province), alpha = 0.2, col = NA) +
#   geom_line(aes(group = Province), alpha = 0.5) +
#   guides(col = "none",fill = "none") +
#   labs(x = NULL, y = "Province-specific long-term trend") -> p1_comb
# 
# p2$plot$labels <- list(x = NULL, y = "Overall seasonal trend")
# p2$plot$scales$scales[[1]] <- scale_x_discrete(limits = month.abb) 
# p2$plot$theme <- theme(axis.text.x = element_text(angle = 45, hjust = 1))
# 
# fit_bym_st_prov %>%
#   gather_draws(r_t[t, Intercept]) %>%
#   median_qi() |> 
#   left_join(select(fitdata, t, month) |> distinct()) -> summ_r_t

```

```{r}
# 
# p1_comb
# p2
# 
# summ_r_t |>  
#   ggplot(aes(month, .value,
#              ymin = .lower, ymax = .upper)) + 
#   geom_hline(yintercept = 0, col = "grey") +
#   geom_errorbar() + 
#   geom_point() +
#   labs(x = "Month", y = "Posterior estimate", subtitle = "Temporal random effect")

```
### Seasonal spline by province {.tabset}

#### Formula and priors
```{r}
# 
# f_bym_prov_seas = bf(
#   n_npev | trials(n_npafp) ~     
#     # Overall average prevalence
#     1 + 
#     # Random deviations by province and time
#     (1|adm1_name) + (1|t) + 
#     # Temporal spline 
#     s(t, k = 10) +
#     # Correlated/uncorrelated random deviations by district 
#     car(W, gr = guid, type = "bym2") + 
#     # Seasonal temporal splines by province
#     s(month_of_year, bs = "cc", k = 12, by = adm1_name) 
#   )

```

#### Fitting

```{r}

# fit_bym_prov_seas <- update(fit_bym_st, 
#                   formula = f_bym_prov_seas,
#                   # Use same priors as before
#                   prior = prior_bym_st,
#                   newdata = fitdata,
#                   data2 = list(W = W),
#                   cores = parallel::detectCores(),
#                   iter = 4000,
#                   control = list(adapt_delta = 0.97),
#                   recompile = T,
#                   file = here(outdir,"fit_bym_prov_seas.rds")) 

```

#### Check fit
```{r}
# summary(fit_bym_prov_seas)
# bayes_R2(fit_bym_prov_seas)
```

```{r}

# fit_bym_prov_seas |> 
#     conditional_smooths() |>
#     plot(plot = F) -> plots
#   
#   ggplot_build(plots[[1]]) -> p1
#   ggplot_build(plots[[2]]) -> p2
# 
# p1$plot$labels <- list(x = NULL, y = "Overall long-term trend")
# 
# p2$plot$data |> 
#   rename(Province = adm1_name) |> 
#   ggplot(mapping = p2$plot$mapping) + 
#   # geom_hline(yintercept = 0, lty = "dashed", col = "darkgrey") +
#   geom_ribbon(aes(group = Province), alpha = 0.2, col = NA) +
#   geom_line(aes(group = Province), alpha = 0.5) +
#   guides(fill = "none") +
#   scale_x_discrete(limits = month.abb) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#   labs(x = NULL, y = "Province-specific seasonal trend", col = "Province") -> p2_comb
# 
# fit_bym_prov_seas %>%
#   gather_draws(r_t[t, Intercept]) %>%
#   median_qi() |> 
#   left_join(select(fitdata, t, month) |> distinct()) -> summ_r_t

```

```{r}

# p1
# p2_comb
# 
# summ_r_t |>  
#   ggplot(aes(month, .value,
#              ymin = .lower, ymax = .upper)) + 
#   geom_hline(yintercept = 0, col = "grey") +
#   geom_errorbar() + 
#   geom_point() +
#   labs(x = "Month", y = "Posterior estimate", subtitle = "Temporal random effect")

```
## Model Comparison

### LOO / WAIC
```{r}

fit_base <- add_criterion(fit_base, c("loo","waic"), moment_match = T)
fit_bym <- add_criterion(fit_bym, c("loo","waic"), moment_match = T)
# Doesn't run
# fit_ar <- add_criterion(fit_ar, c("loo","waic"), moment_match = T, 
#                         recompile = T)
fit_bym_st <- add_criterion(fit_bym_st, c("loo","waic"), moment_match = T)
fit_bym_st_guid <- add_criterion(fit_bym_st_guid, c("loo","waic"), moment_match = T, 
                              file = here("output/fit_bym_st_guid"))
fit_bym_prov_seas <- add_criterion(fit_bym_prov_seas, c("loo","waic"), moment_match = T, 
                              file = here("output/fit_bym_prov_seas"))


models <- list(fit_base, fit_bym, #fit_ar, 
               fit_bym_st,fit_bym_st_guid, fit_bym_prov_seas)

lapply(models, loo)
loo_compare(fit_base, fit_bym, #fit_ar, 
            fit_bym_st, fit_bym_st_guid, fit_bym_prov_seas, 
            criterion = "loo")
loo_compare(fit_base, fit_bym, #fit_ar, 
            fit_bym_st, fit_bym_st_guid, fit_bym_prov_seas,
            criterion = "waic")

```

### PIT
```{r}

lapply(models,pp_check, type = "pit_ecdf")

```
BYM_st_guid seems to be the least skewed, but all within the boundaries. 

### R-squared
```{r}

lapply(models, \(x){bayes_R2(x) |> as.data.frame()}) |>
  setNames(c("Base","BYM", #"BYM_AR", 
             "BYM_st", "BYM_st_guid","BYM_prov_seas")) |> 
  bind_rows(.id = "Model") -> r2_tab

r2_tab
write_csv(r2_tab, "../output/r2_tab.csv")

```
Little difference in R2 -> same amount of variability being explained just in different ways. 

```{r}

r2_tab |> 
  ggplot(aes(y = Model, x = Estimate, xmin = `Q2.5`, xmax = `Q97.5`)) +
  geom_errorbarh(height = 0.5) +
  xlim(0,1) +
  labs(x = "Bayesian R-squared", y = NULL)

ggsave("../figures/compare_r2.png", height = 4, width = 6)
```

### Residuals
```{r}

res <- lapply(models, 
              \(x){residuals(x) |> as.data.frame() |> bind_cols(select(fitdata, n_npafp))}) |> 
  setNames(c("Base","BYM", #"BYM_AR", 
             "BYM_st", "BYM_st_guid","BYM_prov_seas")) |> 
  bind_rows(.id = "Model")
  
```

```{r}

res |> 
  tbl_summary(by = Model) |> 
  add_overall() |> 
  as_gt()

res |> 
  ggplot(aes(x = Estimate, col = Model)) + 
  geom_density() + 
  scale_colour_viridis_d()

res |> 
  ggplot(aes(n_npafp, Estimate)) + 
  geom_jitter(alpha = 0.01) + 
  facet_wrap(~Model) + 
  scale_x_continuous(trans = "sqrt")

```
